If the Null Fits, You Must Omit: Measurement Artifacts Undermine Group Polarization Replications

### 2025-10-31 GroupPolarizationStatmod/paper/method.tex
update method
updated method

Our method combines several formal and computational tools to identify which
replications should be retracted and evaluate experimental designs across our
corpus. The first step in the analysis is to find a reasonable pair of latent
null-polarization distributions that appear polarized due to clipping.  To do
this we designed a web app that allows the user to try different latent
distribution parameters (Figure~\ref{fig:data-model-flow}) to seed a
hillclimbing algorithm that explores latent distribution parameter space
until it finds null parameters that look like group polarization under the
experiment's ordinal measurement procedure, or the algorithm stops after a
certain number of steps (Algorithm~\ref{alg:hillclimbing}).

@@

\begin{figure}
  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
  Latent model participant opinions are drawn from a normal distribution with the
  listed parameters (A,C). The pre-deliberation latent distribution has high
  variance, but appears polarized when measured via integration (A).  The latent
  distribution's variance is expected to decrease during deliberation (B) via
  consensus, but in this example the latent mean is static, 
$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
the simulated observed mean increases, i.e., spurious
group polarization occurs (C).}
  \centering
  \includegraphics[width=0.95\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
  \label{fig:distros}
\end{figure}

>>>>

Our method combines several formal and computational tools to identify which
replications should be retracted and evaluate experimental designs across our
corpus. The first step in the analysis is to find a reasonable pair of latent
null-polarization distributions that appear polarized due to clipping.

\subsection{Identify replications for removal}

The first step is to find which replications of the 57 must be retracted 
due to non-identifiability. We establish non-identifiability by finding one
set of latent null-polarization opinion distributions where the mean doesn't change,
but appear polarized when measured on the ordinal scale used in the experiment.

To do
this we designed a web app that allows the user to try different latent
distribution parameters (Figure~\ref{fig:data-model-flow}) to seed a
hillclimbing algorithm that explores latent distribution parameter space
until it finds null parameters that look like group polarization under the
experiment's ordinal measurement procedure, or the algorithm stops after a
certain number of steps (Algorithm~\ref{alg:hillclimbing}). If the web app
finds

@@

\begin{figure}[ht]
  \caption{\textbf{Opinion measurement model and spurious group polarization.} 
  Latent model participant opinions are drawn from a normal distribution with the
  listed parameters (A,C). The pre-deliberation latent distribution has high
  variance, but appears polarized when measured via integration (A).  The latent
  distribution's variance is expected to decrease during deliberation (B) via
  consensus, but in this example the latent mean is static, 
$\mupre = \mupost$. However, when the post-deliberation distribution is measured,
the simulated observed mean increases, i.e., spurious
group polarization occurs (C).}
  \centering
  \includegraphics[width=0.95\textwidth]{Figures/Model/latent-ordinal-distros.pdf}
  \label{fig:distros}
\end{figure}
